### 2022-08-30 main.tex
initial draft of chapters

\documentclass[twoside,openright,frontopenright]{ip3thesis}
\input{preamble}
\usepackage{subfiles}

\begin{document}
\title{Emulation of high multiplicity matrix elements for electron-positron and hadron-hadron colliders}
% \subtitle{An optional subtitle}
\author{Henry Truong}
\researchgroup{Institute for Particle Physics Phenomenology}
\maketitlepage*

\begin{abstract}
%
	This is some abstract about this thesis.
%
\end{abstract}

\disableprotrusion
\tableofcontents*
\listoffigures
\listoftables
\enableprotrusion

\begin{declaration*}
%
	The work in this thesis is based on research carried out in the Department of
	Physics at Durham University. No part of this thesis has been
	submitted elsewhere for any degree or qualification.
%
\end{declaration*}

\begin{acknowledgements*}
%
	I would like to thank\ldots
%
\end{acknowledgements*}

\begin{epigraph*}
%
	This a place for some profound quote which sums up your work, feelings or
	life.
	\source{Some book}{A. Author}
%
\end{epigraph*}

%\dedicationtext{This thesis is dedicated to}
\begin{dedication*}
%
	Someone special
	\also[and]{that other guy}
%
\end{dedication*}

\cleardoublepage

\subfile{intro}
\subfile{qcd}
\subfile{ml}
\subfile{paper1}
\subfile{paper2}
\subfile{paper3}

\appendix
%\subfile{appendix1}
%\subfile{appendix2}

\bibliographystyle{JHEP}
\bibliography{bibliography}

\end{document}