### 2022-08-30 paper1.tex

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Emulation of tree-level $e^{+}e^{-}$ matrix elements}
\label{chapter:fame1}

\end{document}