### 2022-08-30 paper2.tex

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Emulation of $e^{+}e^{-}$ NLO k-factors}
\label{chapter:fame2}

\end{document}