### 2022-08-30 paper3.tex

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Emulation of hadron-hadron initiated matrix elements}
\label{chapter:fame3}

\end{document}