### 2022-09-21 appendix1.tex
wrote: matrix element factorisation

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Catani-Seymour notation}
\label{appendix:catani_seymour}

\end{document}