### 2022-09-23 qcd.tex

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Quantum chromodynamics in practice}
\label{chapter:qcd}
    In the previous chapter we introduced the basic concepts
    of QCD and setup the framework of calculating
    cross-sections. Cross-sections are the most relevant
    quantities to compute as they can be measured
    experimentally, and are a property of the particles
    being collided, rather than depending on the
    specifics of the experimental procedure.
    This provides a bridge to compare
    theoretical predictions and experiment.
    From (\ref{eqn:dsigma}), the partonic cross-section
    is the matrix element integrated over the relevant
    final-state phase-space, normalised by a flux factor.
    The matrix elements are
    calculated order-by-order in the strong coupling
    $\alpha_{s}$, (\ref{eqn:matrix_element}),
    where each order corresponds to a set of Feynman diagrams
    that have the appropriate number of vertices,
    as each QCD vertex brings along a factor of $\alpha_{s}$.

    In this chapter we will discuss the problems that
    arise when evaluating these Feynman diagrams and outline
    some solutions that have been widely adopted to circumvent
    these issues to provide real-world applicable predictions.
    This chapter forms the theoretical foundations on which
    the work of this thesis is built upon.

    We will consider the divergent nature of matrix
    elements and introduce the most widely used techniques
    to tame these singularities.

\section{Divergent structures}\label{sec:divergences}
    We have seen that matrix element computations
    result in evaluating Feynman diagrams. From
    the QCD Feynman rules (ref here), it becomes
    clear that integrating over all momentum modes
    will give rise to singularities. The divergences
    associated with high energy modes are ultraviolet
    (UV) divergences. On the opposite end of the energy
    spectrum, low energy modes also give rise to 
    infrared (IR) divergences.

    The de-facto methods to alleviate these
    divergences are through regularisation
    for UV divergences, and through subtraction
    schemes for IR divergences. Both of these
    methods will be discussed in this section.

\subsection{Ultraviolet divergences}\label{sec:UV}
    During intermediate steps of calculations,
    such as trying to compute loop diagrams seen in
    the Figure~\ref{fig:bubble_diagram},
    we have to evaluate integrals of the form
    \begin{equation}\label{eqn:loop_integral}
        I_{\mathrm{UV}} = \int_{0}^{\Lambda} \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} \sim \log{\Lambda} \, ,
    \end{equation}
    where a cutoff scale $\Lambda$ has been introduced to
    capture the divergence as the loop momenta
    $k \rightarrow \infty$. It is clear that the
    integral diverges in this high energy limit,
    hence the name ultraviolet divergence.
    This UV divergence cannot influence the
    experimentally observable quantities at experiments,
    which are finite, because the final observable
    should not depend on the intermediate steps taken
    to get there.
    Infinities associated with UV divergences
    are removed via the process of renormalisation
    which introduces counterterms in the Lagrangian
    to exactly cancel these divergences.

    \begin{figure}
        \includegraphics[scale=1.0]{qcd/bubble_diagram.pdf}
        \caption{Bubble diagram}
        \label{fig:bubble_diagram}
    \end{figure}

    Before we introduce these counterterms, we
    first consider the procedure of regularisation
    which makes these infinities explicitly manifest.
    The most widely adopted method of regularisation
    is dimensional regularisation (DR) \cite{tHooft:1972tcz} which
    is based on the observation that the integral
    (\ref{eqn:loop_integral}), which is carried
    out in $d=4$ space-time dimensions, would be finite
    if $d < 4$.

    In DR we define $d = 4 - 2\epsilon$ such that
    after Feynman parametrisation (\ref{eqn:loop_integral})
    becomes
    \begin{equation}\label{eqn:dim_reg}
        % \int \dfrac{\mathrm{d}^{d}\ell}{(2\pi)^{d}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} = \dfrac{\mathrm{i}(-p^{2})^{-\epsilon}}{(4\pi)^{2-\epsilon}}\dfrac{\Gamma(1-\epsilon)^{2}}{\Gamma(2-2\epsilon)}\Gamma(\epsilon) \, ,
        I_{\mathrm{UV}} = \int \dfrac{\mathrm{d}^{d}\ell}{(2\pi)^{d}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} = \dfrac{\mathrm{i}}{(4\pi)^{2}}\left(\dfrac{-p^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1-\epsilon)^{2}}{\Gamma(2-2\epsilon)}\Gamma(\epsilon) \, ,
    \end{equation}
    where the divergence for $d = 4$, or equivalently
    $\epsilon \rightarrow 0$, is now captured in the
    Gamma function, $\Gamma(\epsilon)$. This becomes clear once
    we expand $\Gamma(\epsilon)$ around small $\epsilon$
    \begin{equation}\label{eqn:gamma_laurent}
        \Gamma(\epsilon) = \dfrac{1}{\epsilon} - \gamma_{E} + \dfrac{1}{2}\left(\gamma_{E}^{2} + \dfrac{\pi^{2}}{6}\right)\epsilon + \mathcal{O}(\epsilon^{2}) \, ,
    \end{equation}
    where $\gamma_{E} \approx 0.577$ is the Euler-Mascheroni
    constant, meaning the divergence is now regularised as a pole in
    $\epsilon$.

\subsection{Renormalisation}\label{sec:renormalisation}
    With the UV divergence regularised by dimensional
    regularisation, it is now possible to construct the forms
    of the counterterms order-by-order in $\alpha_{s}$ to
    explicitly cancel them.

    A subtlety with adjusting the dimension of the theory
    from $d=4$ to arbitrary dimensions is that the mass dimension
    of the Lagrangian has to change accordingly to retain a
    dimensionless action. We can account for this by
    introducing an arbitrary scale $\mu_{R}$, the renormalisation
    scale, into the gauge coupling
    constant to modify the mass dimension of the Lagrangian.
    For QCD this is done via the modification
    \begin{equation}\label{eqn:dimensionful_coupling}
        g_{s} \rightarrow g_{s}\mu_{R}^{\epsilon} \, .
    \end{equation}
    where it is understood the original gauge coupling is
    dimensionless. The order of $\mu_{R}$ is determined by examining
    the dimensions of the gauge and fermion fields
    in the Lagrangian (\ref{eqn:L_QCD}).
    We see that this introduces a scale dependence
    on the coupling constant $\alpha_{s}$, a point we will return
    to in \ref{sec:alpha_running}.
    % In a calculation,
    % intermediate steps will have a dependence on $\mu$,
    % however, once the $\epsilon \rightarrow 0$ limit
    % is taken the scale dependence drops out.

    The inclusion of counterterms into the theory can be
    thought of as renormalisations of the the bare fields
    and couplings in (\ref{eqn:L_QCD}) to restore predictive
    power to the theory. The renormalised fields and couplings
    can be given as (see e.g. Ref.~\cite{mangano1999introduction})
    \begin{align}\label{eqn:renormlised_fields}
        \psi_{\mathrm{bare}} &= \sqrt{Z_{2}} \psi_{R} \, , \nonumber \\
        A_{\mathrm{bare}}^{\mu} &= \sqrt{Z_{3}} A_{R}^{\mu} \, ,\\
        g_{s,\mathrm{bare}} &= Z_{g} \mu_{R}^{\epsilon} g_{s,R} \, , \nonumber
    \end{align}
    where it conventional to define $Z_{1} = Z_{g}Z_{2}\sqrt{Z_{3}}$
    such that we can set $Z_{n} = 1 + \delta_{n}$ for $n \in \{1, 2, 3\}$.
    In this way we can write the Lagrangian as
    \begin{equation}
        \mathcal{L}_{\mathrm{renorm}} = \mathcal{L}_{\mathrm{bare}} + \mathcal{L}_{\mathrm{c.t.}} \, ,
    \end{equation}
    where the counterterms appearing in $\mathcal{L}_{\mathrm{c.t.}}$
    are determined by calculating $\delta_{n}$. There is freedom
    in the choice of $\delta_{n}$ which gives rise to
    different regularisation schemes. The minimal subtraction
    (MS) scheme subtracts only the epsilon pole appearing
    in loop integrals (\ref{eqn:gamma_laurent}). However, the most
    commonly used regularisation scheme is the modified minimal
    subtraction ($\overline{\mathrm{MS}}$) scheme which subtracts the
    epsilon pole plus a universal constant appearing in all
    loop integrals.

    In an all orders calculation of an observable,
    there would be no dependence on the renormalisation
    scale as it is a remnant of the regularisation prescription
    However, because in perturbative QCD observables
    are calculated order-by-order, there will be a residual
    dependence on the renormalisation scale stemming from
    the missing higher order terms. The conventional way
    that the community estimates the uncertainty associated
    with this residual dependence is by carrying out
    a scale variation. This involves evaluating the
    relevant quantities at the nominal renormalisation
    scale and again with the scale varied by a factor of
    two above and below.

\subsection{Running of the coupling constant}\label{sec:alpha_running}
    The bare coupling should not depend on the actual
    value of the renormalisation scale,
    \begin{equation}\label{eqn:d_g_bare_d_mu}
        \dfrac{\mathrm{d}g_{\mathrm{bare}}}{\mathrm{d}\mu_{R}} = 0 \, ,
    \end{equation}
    or said another way, the renormalisation process
    is independent of the renormalisation scale. The
    consequence of this is that $\alpha_{s}$ has to
    depends on the renormalisation scale.

    The behaviour of $\alpha_{s}$
    is governed by the Callan-Symanzik \cite{Callan:1970yg,Symanzik:1970rt}
    $\beta$-function
    \begin{equation}\label{eqn:beta_fn}
        \mu^{2} \dfrac{\partial \alpha_{s}(\mu^{2})}{\partial \mu^{2}} = \beta(\alpha_{s}) \, ,
    \end{equation}
    where the $\beta$-function can be written as a perturbative
    expansion in $\alpha_{s}$
    \begin{equation}\label{eqn:beta_series}
        -\beta(\alpha_{s}) = \alpha_{s}\sum_{n=0}^{\infty}\left(\dfrac{\alpha_{s}}{4\pi}\right)^{n+1} \beta_{n}\, 
    \end{equation}
    where the coefficients $\beta_{n}$ have been computed
    up to $\beta_{4}$ \cite{Baikov:2016tgj,Luthe:2017ttg}.
    The solution to (\ref{eqn:beta_fn}) to first order
    is
    \begin{equation}\label{eqn:1l_alpha}
        \alpha_{s}(\mu^{2}) = \dfrac{1}{\frac{\beta_{0}}{4\pi}\log\left(\frac{\mu^{2}}{\Lambda_{\mathrm{QCD}}^{2}}\right)} \, ,
    \end{equation}
    where $\Lambda_{\mathrm{QCD}}$ is the QCD scale,
    the scale at which $\alpha_{s}$ becomes
    large enough that perturbation theory breaks down,
    and $\beta_{0} = (11C_{A} - 4T_{R}n_{f})/3$.
    In QCD, $C_{A}=3$ and $T_{R}=\frac{1}{2}$, meaning for
    $n_{f} < \frac{33}{2}$ the sign of $\beta_{0}$ is positive,
    which is what we have observed in nature with six
    flavours of quarks. This behaviour of positive
    $\beta_{n}$ is also the case in all coefficients
    computed to date, meaning that $\alpha_{s}$
    decreases with increasing energy due to the minus sign
    in (\ref{eqn:beta_series}). This property is known as
    asymptotic freedom \cite{Gross:1973id,Politzer:1973fx},
    and is justification for treating QCD perturbatively.
    The running of the coupling constant has been
    observed experimentally as illustrated in Figure~\ref{fig:alpha_s_running}.

    \begin{figure}
        \includegraphics{qcd/alpha_s_running.pdf}
        \caption{The running of the strong coupling constant $\alpha_{s}$,
        as determined by experiments, with QCD theory prediction in black.
        Figure reference \cite{Workman:2022ynf}.}
        \label{fig:alpha_s_running}
    \end{figure}

    Another consequence of the running coupling
    is colour confinement, in which quarks and
    gluons cannot exist as free particles. Due
    to the increasing strength of coupling at low
    energies, quarks and gluons are forced to
    form composite, colourless particles. This means
    that at low energies, perturbative QCD
    is not an accurate description of nature,
    and the behaviour of hadrons cannot be predicted
    from perturbation theory.

\subsection{Infrared divergences}
    In Section~\ref{sec:UV} we saw how divergences
    stemming from high energy behaviour would occur
    in loop diagrams. On the other end of the energy
    spectrum we also have divergences from low energy
    modes. Since these divergences emerge from low energy
    behaviour they are called infrared (IR) divergences. There
    are two cases in which these divergences arise:
    in loop integrals (virtual) and when an emission
    of an extra particle (c.f. splitting functions
    (\ref{eqn:AP_kernels})) has vanishing energy (soft),
    or becomes parallel to an external colourful line (collinear).

    Virtual IR divergences can be understood by looking
    at an integral of the form
    \begin{equation}\label{eqn:virtual_IR_integral}
        I_{v} = \int \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}} \dfrac{1}{\ell^{2}(\ell+p_{1})^{2}(\ell-p_{2})^{2}} \, ,
    \end{equation}
    which is encountered when calculating the one-loop
    virtual correction to the gluon-quark-antiquark vertex
    as shown in Figure~\ref{fig:vertex_correction}.
    It is clear that the denominator vanishes when $\ell \rightarrow 0$
    or when either $(\ell + p_{1})^{2}$ or $(\ell - p_{2})^{2} \rightarrow 0$.
    These situations correspond to the gluon in the loop propagator
    going soft, or collinear to the external quark or antiquark.
    These divergences are regulated in DR to give
    \begin{equation}\label{eqn:virtual_IR_DR}
        I_{v} = \dfrac{\mathrm{i}}{(4\pi)^{2}Q^{2}}\left(\dfrac{-Q^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1+\epsilon)\Gamma(1-\epsilon)^{2}}{\Gamma(1-2\epsilon)}\left[\dfrac{1}{\epsilon^{2}}\right] \, ,
    \end{equation}
    where $Q = (p_{1} + p_{2})$. There is a double $\epsilon$ pole manifest,
    corresponding to the soft and collinear divergence.

    \begin{figure}
        \includegraphics[scale=1.0]{qcd/vertex_correction.pdf}
        \caption{Vertex correction}
        \label{fig:vertex_correction}
    \end{figure}

    Real IR divergences are manifest when integrating
    matrix elements over the phase-space of external state momenta.
    Matrix elements can be written as functions of Mandelstam variables
    \begin{equation}\label{eqn:mandelstam}
        s_{ij} = (p_{i} + p_{j})^{2} \xrightarrow[\mathrm{limit}]{\mathrm{massless}} 2 E_{i}E_{j}(1-\cos \theta_{ij}) \, ,
    \end{equation}
    where $p_{i}$ and $p_{j}$ are the 4-momenta of partons $i$ and $j$
    in the hard scattering.
    When either of the partons beecome soft, $E_{i,j}\rightarrow 0$,
    or when they go collinear to each other, $\theta_{ij} \rightarrow 0$,
    the matrix element will diverge if $s_{ij}$ appears
    in the demoninator. These scenarios correspond to an external
    particle becoming unresolved.

    In Section~\ref{sec:renormalisation}, we removed UV divergences
    through the use of counterterms to renormalise the theory, making
    predictions from the theory physical.
    IR divergences, on the other hand, arise even after the theory
    has been renormalised. The resolution to IR divergences is given by the
    Block-Nordsieck \cite{Bloch:1937pw,Yennie:1961ad} and Kinoshita-Lee-Nauenberg
    (KLN) theorems \cite{Kinoshita:1962ur,Lee:1964is} which states
    that at each order of perturbation theory, the virtual and real
    contributions exactly cancel out, leaving only finite corrections.
    This cancellation occurs because the infrared pole structure is
    identical in the real and virtual corrections, but with an opposite sign.

    At a detector of a collider experiment, there is a finite
    energy resolution of the calorimeter. This means that
    it is not physically possible to observe arbitrarily soft
    partons. It is also not possible to distinguish two partons
    at arbitrarily small angles from one parton with the combined
    momenta. These limitations along with the KLN theorem
    means that any physical observable that we wish to predict
    with fixed-order perturbation theory must be insensitive to the
    emissions of soft or collinear partons. Any observable obeying
    this criteria is an infrared and collinear (IRC) safe observable.
    This leads to the concept of jets \cite{Salam:2010nqg} which are collimated
    particles combined according to some jet algorithm.

    In DR, the singularities from soft and collinear divergences
    are manifest as $\epsilon$ poles which makes the cancellation
    simple between the real and virtual parts. However, in practice, 
    the phase-space integrals of the matrix elements are rarely carried
    out analytically due to the high dimensionality of the integral.
    Instead they are done numerically with the use
    of Monte Carlo methods, which requires the integral to be in
    integer dimensions. Therefore, numerical techniques
    have been developed to deal with these IR divergences as well. The
    most common technique, subtraction, will be discussed in
    more detail in Section~\ref{sec:subtraction}.

\section{Factorisation of matrix elements}\label{sec:me_factorisation}
    In the previous section we examined the IR divergences
    that can arise in matrix elements, either explicitly
    through loop diagrams, or implicitly when carrying
    out the phase-space integral over the external state
    momenta. In this section we will discuss how these IR
    singularities are universal and how the divergences
    associated with real emissions can be factorised out
    of the matrix element. This property is exploited
    extensively in Chapters~\ref{chapter:fame1}, \ref{chapter:fame2},
    and \ref{chapter:fame3}
    where the research of this thesis is presented.

    The regions of phase-space in which real emission
    matrix elements diverge are when the emission is
    soft and/or collinear. It can be shown that in these
    regions of phase-space, the matrix element factorises
    into a process-independent singular factor, multiplied
    by a matrix element with the unresolved parton
    being absorbed by an emitting leg (reduced matrix
    element). This factorisation, however, is not exact
    in QCD where there are spin and colour correlations.

    Schematically, the $(n+1)$-body matrix element
    factorises as
    \begin{equation}\label{eqn:me_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow \bm{S}_{ijk} \otimes |\mathcal{M}_{n}|^{2} \, ,
    \end{equation}
    where $\bm{S}_{ijk}$ is a universal singular factor capturing
    the IR divergent behaviour and $|\mathcal{M}_{n}|^{2}$
    is the reduced matrix element. The $\otimes$
    represents the colour and spin correlations existing
    between the singular function and the reduced matrix element.
    This singular factor
    is not unique and can be represented by different
    approximations as long as they reproduce the
    correct IR behaviour. In this thesis we discuss
    two approximations: Catani-Seymour dipoles \cite{Catani:1996vz}
    in Section~\ref{sec:CS_dipoles} and antenna functions \cite{Gehrmann-DeRidder:2005btv}
    in Section~\ref{sec:antenna_functions}. The indices
    of $\bm{S}_{ijk}$ already hints at the fact that the singular
    functions only depends on three partons, and not all
    of the $n+1$ external states.

    We will inspect the soft and collinear limits separately
    to see how the matrix element factorises in these respective
    limits.
    The following section will follow the conventions
    of Ref.~\cite{Catani:1996vz}, namely that the colour and
    helicity summed $n$-body matrix element written as
    \begin{equation}\label{eqn:CS_matrix_element}
        |\mathcal{M}_{n}|^{2} = {}_{n}\langle 1, \ldots, n | 1, \ldots, n \rangle_{n}
    \end{equation}
    where $| 1, \ldots, n \rangle_{n}$ is a vector
    in colour + helicity space (see Appendix~\ref{appendix:catani_seymour}
    for a more thorough explanation on notation).

\subsection{Soft limits}\label{sec:me_soft}
    Consider a tree-level matrix element $|\mathcal{M}_{n+1}|^{2}$
    with a final-state gluon $j$. Note that reduced matrix
    elements associated with taking a quark soft have to vanish
    due to violation of quark number, therefore only gluons
    are considered.
    The limit of the soft gluon with momentum
    $p_{j}$ can be parametrised by
    \begin{equation}\label{eqn:soft_parametrisation}
        p_{j}^{\mu} = \lambda q^{\mu} \, , \quad \lambda \rightarrow 0 \, ,
    \end{equation}
    where $q^{\mu}$ is an arbitrary four-vector and $\lambda$
    is a scale parameter. In this limit the matrix element
    can be written as
    \begin{equation}\label{eqn:soft_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow -\dfrac{1}{\lambda^{2}}8\pi \mu^{2\epsilon}\alpha_{s} \sum_{i}\dfrac{1}{p_{i}q}\sum_{k \neq i} \dfrac{p_{k}p_{i}}{(p_{i} + p_{k})q} \; {}_{n}\langle 1, \ldots, n+1 | \bm{T}_{k} \cdot \bm{T}_{i} | 1, \ldots, n+1 \rangle_{n} \, ,
    \end{equation}
    where terms less singular than $1 / \lambda^{2}$ have been
    neglected. The reduced matrix element is obtained by removing
    the soft gluon. The indices $i$, $j$, and $k$ label partons involved
    in the factorsation process: $i$ is the emitter parton,
    $j$ is the emitted parton, and $k$ is a parton accounting
    for the colour correlations. The scale $\mu$ can be identified as the
    renormalisation scale, and $\bm{T_{i}}$ is the colour-charge
    operator. It acts on the colour vectors to give factors of
    $C_{A}$ and $C_{F}$, hence this factorisation is not exact.

    The matrix elements in (\ref{eqn:soft_factorisation}) are
    unambiguously defined only when momentum conservation is
    fulfilled. This is only true in the strict $\lambda = 0$
    limit. Away from the limit, care has to be taken to conserve
    momentum conservation and keep the relevant partons on-shell.
    This can be done through the use of momentum mappings
    \cite{Catani:1996vz,Kosower:1997zr}.

\subsection{Collinear limits}\label{sec:me_collinear}
    For the collinear limit, consider partons $i$ and
    $j$ in the matrix element $|\mathcal{M}_{n+1}|^{2}$.
    Their momenta can be decomposed as
    \begin{align}\label{eqn:F_collinear_momenta}
        \begin{split}
        p_{i}^{\mu} &= z p^{\mu} + k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{z} \dfrac{n^{\mu}}{2 p n} \, \\
        p_{j}^{\mu} &= (1-z) p^{\mu} - k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{1-z}\dfrac{n^{\mu}}{2 p n} \, , 
        \end{split}
    \end{align}
    where $p^{\mu}$ denotes the collinear direction
    of the two partons, $k_{\perp}^{\mu}$ specifies
    the transverse direction perpendicular to the
    collinear direciton ($p k_{\perp} = 0$), and $n^{\mu}$ is an auxiliary
    vector satisfying the conditions $n^{2} = 0$ and $k_{\perp} n = 0$.
    $z$ is the fraction of momenta carried away from
    the collinear momentum by parton $i$.

    The collinear limit can be then be defined as
    \begin{equation}\label{eqn:F_collinear limit}
        2 p_{i} p_{j} = s_{ij} = - \dfrac{k_{\perp}^{2}}{z(1-z)} \, , \quad k_{\perp} \rightarrow 0 \, .
    \end{equation}
    In this limit the matrix element can be written as
    \begin{equation}\label{eqn:F_collinear_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow \dfrac{1}{p_{i}p_{j}}4\pi \mu^{2\epsilon}\alpha_{s} \; {}_{n} \langle 1, \ldots, n+1 | \hat{P}_{ij}(z, k_{\perp}) | 1, \ldots, n+1 \rangle_{n} \, ,
    \end{equation}
    where terms less singular than $k_{\perp}^{2}$ have been neglected.
    This reduced matrix element is obtained by replacing the partons
    $i$ and $j$ with a single parton $ij$ which carries the
    momentum $p^{\mu}$ and suitable quantum numbers depending
    on the partons $i$ and $j$. For example, if $i =$ quark and $j=$ gluon,
    then $ij =$ quark, or if $i=$ quark and $j =$ antiquark, then $ij =$ gluon.
    $\hat{P}_{ij}$ are the $d$-dimensional Altarelli-Parisi splitting
    functions that depend on the momentum fraction $z$ for the splitting
    $ij \rightarrow i + j$ and the transverse momentum $k_{\perp}$.
    Each splitting function is matrix acting on the spin indices of $ij$.
    Due to these spin correlations the matrix
    element does not factorise from the splitting functions exactly.

    $\hat{P}_{ij}$ become the more recognisable Altarelli-Parisi
    splitting functions (\ref{eqn:AP_kernels}) once spin-averaged
    and the $\epsilon \rightarrow 0$ limit is taken.

    In general, it is possible for the case in which a final-state
    parton $i$ becomes collinear with an initial-state parton $a$.
    This is described by the splitting process $a \rightarrow ai + i$.
    In this case their momenta can be decomposed as
    \begin{align}\label{eqn:I_collinear_momenta}
        \begin{split}
        p_{i}^{\mu} &= (1-x)p_{a}^{\mu} + k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{1-x}\dfrac{n^{\mu}}{2p_{a} n} \, , \\
        p_{a}^{\mu} &= x p_{a}^{\mu} \, ,
        \end{split}
    \end{align}
    with the collinear limit defined as
    \begin{equation}\label{eqn:I_collinear_limit}
        2p_{i}p_{a} = s_{ia} = -\dfrac{k_{\perp}^{2}}{1-x} \, , \quad k_{\perp} \rightarrow 0 \, .
    \end{equation}

    The analogous expression of (\ref{eqn:F_collinear_factorisation})
    for initial-state splitting is
    \begin{equation}\label{eqn:I_collinear_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow \dfrac{1}{x} \dfrac{1}{p_{i}p_{a}} 4\pi \mu^{2\epsilon} \alpha_{s} \; {}_{n} \langle 1,\ldots,n+1;ai, \ldots | \hat{P}_{ai}(x, k_{\perp}) | 1,\ldots,n+1;ai,\ldots \rangle_{n} \, ,
    \end{equation}
    where the replacement of partons $a$ and $i$ have been made
    explicit by the presence of the parton $ai$ in the colour + helicity vector.
    The specific type of parton $ai$ depends upon the types of $a$ and $i$.

    Similar to (\ref{eqn:soft_factorisation}), (\ref{eqn:F_collinear_factorisation})
     and (\ref{eqn:I_collinear_factorisation}) are only
    true in the strict collinear limit. Away from these
    limits care has to be taken to conserve momenta
    via the use of momenta mappings.

\subsection{Factorisation of colour-ordered amplitudes}
    The factorisation expressed in (\ref{eqn:soft_factorisation}),
    (\ref{eqn:F_collinear_factorisation}), and (\ref{eqn:I_collinear_factorisation})
    were not exact because of colour and spin correlations.

    It can be shown that the factorisation of matrix
    elements in the soft and collinear limits becomes
    exact once the colour structure of the gauge group
    is separated from the kinematics.

    For the scope of this thesis it is sufficient
    to consider colour decomposition of electron-positron
    annihilation. In general any QCD amplitude can be
    colour decomposed. Consider the process $e^{+}e^{-} \rightarrow q\bar{q} + n g$,
    the amplitude can be written as a product of hadronic and
    leptonic currents \cite{Campbell:1998nn} (on which this discussion is based on)
    \begin{equation}\label{eqn:epem_amplitude}
        \mathcal{M}(q_{1}, \bar{q}_{2}; 1, \ldots, n) = \hat{\mathcal{S}}_{\mu}^{n+2}(q_{1}; 1,\ldots,n; \bar{q}_{2})V^{\mu} \, ,
    \end{equation}
    where the leptonic current is $V^{\mu}$ and the hadronic
    current is
    \begin{equation}\label{eqn:hadronic_current}
        \hat{\mathcal{S}}_{\mu}^{n+2}(q_{1};1,\ldots,n;\bar{q}_{2}) = \mathrm{i}eg_{s}^{n} \sum_{P(1,\ldots,n)}(T^{a_{1}} \ldots T^{a_{n}})_{c_{1}c_{2}}S_{\mu}(q_{1};1,\ldots,n;\bar{q}_{2}) \, .
    \end{equation}
    In this expression we have the electromagnetic gauge coupling
    $e$ and the strong gauge coupling $g_{s}$ appearing. The colour
    structure has been factored into a product of fundamental group
    generators where the indices $a_{i} \in {1,\ldots, N_{c}^{2}-1}$
    and $c_{i} \in {1,\ldots,N_{c}}$. This leaves the colour-ordered
    partial amplitude $S_{\mu}(q_{1};1,\ldots,n;\bar{q}_{2})$ depending
    only on kinematical variables, and where the gluons are emitted in an
    ordered fashion from the quarks. The sum over $P(1,\ldots,n)$
    represents a sum over all permutations of gluon emissions
    which accounts for all Feynman diagrams and colour structures.

    Upon squaring the amplitude (\ref{eqn:epem_amplitude}), we get
    \begin{equation}\label{eqn:epem_me2}
        \left|\hat{\mathcal{S}}_{\mu}^{n+2}V^{\mu}\right|^{2} = e^{2}\left(\dfrac{g_{s}^{2}N}{2}\right)^{n}\left(\dfrac{N_{c}^{2}-1}{N_{c}}\right) \sum_{P(1,\ldots,n)} \left(\left|S_{\mu}(q_{1};1,\ldots,n;\bar{q}_{2})V^{\mu}\right|^{2} + \mathcal{O}\left(\dfrac{1}{N_{c}^{2}}\right)\right) \, ,
    \end{equation}
    where the subleading colour terms have been omitted.

    With the amplitude written in terms of the colour-ordered
    partial amplitudes, it is now possible to factorise (\ref{eqn:epem_me2})
    exactly in the soft and collinear limits. In the limit where gluon
    $j$ is soft, we have
    \begin{equation}\label{eqn:CO_soft_factorisation}
        \left|S_{\mu}(q_{1};1,\ldots,i,j,k,\ldots,n;\bar{q}_{2})V^{\mu}\right|^{2} \rightarrow S_{ijk} \left|S_{\mu}(q_{1};1,\ldots,i,k,\ldots,n;\bar{q}_{2})V^{\mu}\right|^{2} \, ,
    \end{equation}
    where the factor
    \begin{equation}\label{eqn:eikonal_factor}
        S_{ijk} = 4\dfrac{s_{ik}}{s_{ij}s_{jk}} \, ,
    \end{equation}
    is the well-known eikonal factor. We see that the colour-ordered
    amplitude on the RHS of (\ref{eqn:eikonal_factor}) has gluon $j$
    removed but the ordering of all hard partons remain unchanged.

    In the collinear limit where partons $i$ and $j$ become collinear
    to form parton $k$, the matrix element factorises as
    \begin{equation}\label{eqn:CO_collinear_factorisation}
        \left|S_{\mu}(q_{1};1,\ldots,i,j,\ldots,n;\bar{q}_{2})V^{\mu}\right|^{2} \rightarrow \dfrac{2}{s_{ij}}P_{ij}(z) \left|S_{\mu}(q_{1};1,\ldots,k,\ldots,n;\bar{q}_{2})V^{\mu}\right|^{2} \, ,
    \end{equation}
    where $P_{ij}(z)$ are the spin-averaged Altarelli-Parisi splitting
    functions (\ref{eqn:AP_kernels}) with the colour factors removed.
    In this expression it is understood that the indices $i$, $j$, and $k$
    have to be self-consistent for the different partonic splittings.
    For partons which are not colour connected (are not neighbouring
    partons in the colour ordered amplitude), there will be no singular
    behaviour as $s_{ij} \rightarrow 0$.

    In (\ref{eqn:CO_soft_factorisation}) and (\ref{eqn:CO_collinear_factorisation})
    the colour-ordered amplitude factorises exactly into a
    colour-ordered amplitude with one parton removed, multiplied
    by a universal singular factor. This singular factor depends on the unresolved
    parton and the two neighbouring hard particles. Interpreting this as
    the two hard particles forming an antenna which radiates the unresolved
    parton gives rise to the antenna functions which will be discussed
    further in Section~\ref{sec:antenna_functions}.

\subsection{One-loop matrix element factorisation}\label{sec:OL_factorisation}
    The discussion of matrix element factorisation so
    far has been focused on tree-level matrix elements.
    It has been shown that one-loop colour-ordered amplitudes
    also factorise in the soft and collinear limits. At the one-loop level,
    there are new universal singular functions and the
    factorisation formula are modified.

    In the soft limit, a one-loop colour-ordered amplitude
    factorises as (cite here)
    \begin{equation}\label{eqn:1L_soft_factorisation}
        M_{n+1}^{(1)}(\ldots, i, j, k, \ldots) \rightarrow S_{ijk}^{(0)} \, M_{n}^{1}(\ldots, i, k, \ldots) + S_{ijk}^{(1)}(\epsilon) \, M_{n}^{0}(\ldots, i, k, \ldots) \, ,
    \end{equation}
    where $S_{ijk}^{(0)}$ is the eikonal factor (\ref{eqn:eikonal_factor}),
    and $S_{ijk}^{(1)}(\epsilon)$ is the one-loop soft radiation function (cite here).

    A similar factorisation formula for the collinear limit is
    \begin{equation}\label{eqn:1L_collinear_factorisation}
        M_{n+1}^{(1)}(\ldots, i, j, \ldots) \rightarrow \dfrac{1}{s_{ij}} \left[ P_{ij}^{(0)}(z) \, M_{n}^{(1)}(\ldots, k, \ldots) + P_{ij}^{(1)}(z, \epsilon) \, M_{n}^{(0)}(\ldots, k, \ldots) \right] \, ,
    \end{equation}
    where $P_{ij}^{(0)}$ are the tree-level splitting functions
    (\ref{eqn:AP_kernels}) and $P_{ij}^{(1)}$
    are the one-loop splitting functions (cite here).

    In both these formulae the structure is of the form:
    tree-level splitting function multipled by a one-loop
    amplitude, plus a one-loop splitting function multiplied
    by a tree-level amplitude as diagrammatically shown
    in Fig ...

    In Section~\ref{sec:antenna_functions} we will give
    a brief overview of antenna functions where these one-loop
    factorisation formulae were used to obtain universal singular functions
    at the one-loop level for squared matrix elements. These
    antenna functions are then applied in the context of NLO
    k-factor emulation in Chapter~\ref{chapter:fame2}.

\section{Subtraction}\label{sec:subtraction}
\subsection{Catani-Seymour dipoles}\label{sec:CS_dipoles}
\subsection{Antenna}\label{sec:antenna_functions}
\end{document}