### 2022-11-02 main.tex

\documentclass[twoside,openright,frontopenright]{ip3thesis}
\input{preamble}
\usepackage{subfiles}

\begin{document}
\title{Emulation of high multiplicity matrix elements for electron-positron and hadron-hadron colliders}
% \subtitle{An optional subtitle}
\author{Henry Truong}
\researchgroup{Institute for Particle Physics Phenomenology}
\maketitlepage*

\begin{abstract}
%
	This is some abstract about this thesis.
%
\end{abstract}

\disableprotrusion
\tableofcontents*
\listoffigures
\listoftables
\enableprotrusion

\begin{declaration*}
%
	The work in this thesis is based on research carried out in the Department of
	Physics at Durham University. No part of this thesis has been
	submitted elsewhere for any degree or qualification.

	Research presented in this thesis are based on joint work
	published in the articles:
	\begin{itemize}
		\item Chapter \ref{chapter:fame1} is based on \cite{Maitre:2021uaa}: Daniel Ma\^{i}tre and Henry Truong, \textit{A factorisation-aware Matrix element emulator}, Journal of High Energy Physics, 2021 (11). 066.
		\item Chapter \ref{chapter:fame2} is based on ...
		\item Chapter \ref{chapter:fame3} is based on ...
	\end{itemize}

	Other research projects carried out during my studies but not included
	in this thesis:
	\begin{itemize}
		\item \cite{doi:10.1098/rsos.210506}: J. Aylett-Bullock, C. Cuesta-Lazaro, A. Quera-Bofarull, M. Icaza-Lizaola, A. Sedgewick, H. Truong, A. Curran, E. Elliott, T. Caulfield, K. Fong, I. Vernon, J. Williams, R. Bower and F. Krauss, \textit{June: open-source individual-based epidemiology simulation}, (2021), R. Soc. open sci.8210506210506.
		\item \cite{doi:10.1098/rsta.2022.0039}: I. Vernon, J. Owen, J. Aylett-Bullock, C. Cuesta-Lazaro, J. Frawley, A. Quera-Bofarull, A. Sedgewick, D. Shi, H. Truong, M. Turner, J. Walker, T. Caulfield, K. Fong and F. Krauss, \textit{Bayesian emulation and history matching of JUNE}, (2022), Phil. Trans. R. Soc. A.3802022003920220039.
	\end{itemize}

%
\end{declaration*}

\begin{acknowledgements*}
%
	I would like to thank\ldots
%
\end{acknowledgements*}

\begin{epigraph*}
%
	This a place for some profound quote which sums up your work, feelings or
	life.
	\source{Some book}{A. Author}
%
\end{epigraph*}

%\dedicationtext{This thesis is dedicated to}
\begin{dedication*}
%
	Someone special
	\also[and]{that other guy}
%
\end{dedication*}

\cleardoublepage

\subfile{intro}
\subfile{qcd}
\subfile{MC}
\subfile{ml}
\subfile{paper1}
\subfile{paper2}
\subfile{paper3}
\subfile{conclusion}

\appendix
\subfile{appendix1}
\subfile{fame_appendex1}
%\subfile{appendix2}

\bibliographystyle{JHEP}
\bibliography{bibliography}

\end{document}